Kindly Bent to Free Us

### 2019-07-08 kindly/infer/infer.tex
Reformulation/mv/fixes.

An important innovation of \affe is its principal type inference.
We now formulate our inference algorithm
based on the HM(X) framework~\citep{DBLP:journals/tapos/OderskySW99}.
HM(X) presents how to implement Hindley-Milner type inference in a language
with qualified types and an arbitrary custom constraint language.
We adapt and extend the rules of HM(X) to support kind inference,
track linearity and handle borrows and regions. Furthermore, we
formulate an appropriate constraint solving algorithm for our constraint
system such that inference is sound and complete.
In addition, our constraint solving algorithm simplifies constraints in a
principled way.
We first present various preliminary definitions
before showing our kind and type inference judgements
and our constraint solving algorithm.

@@

\subsubsection{Constraint desugaring}

\TODO{Present this if there is space}

\begin{figure}[!htb]
  \begin{mathpar}

  \inferrule{}{ \Cleq{\Eempty}{k} \Crewrite  \Cempty}

  \inferrule{
    \Cleq\E k \Crewrite  C \\ \E \vdash \Cleq B k \Crewrite  D
  }{
    \Cleq{\E; B}{k} \Crewrite  C \Cand D
  }

  \inferrule{
    \E \vdash \Cleq \schm k \Crewrite  C
  }{ \E \vdash \Cleq{\bvar x \schm}{k} \Crewrite  C}

  \inferrule{ }{
    \E \vdash \Cleq{\svar x \schm^n} k \Crewrite  \Cfalse
  }

  % \inferrule{}{\Cleq{\IBORROW^n} k \Crewrite  \Cleq{\kun_n} k}

  % \inferrule{}{\Cleq{\MBORROW^n} k \Crewrite  \Cleq{\kaff_n} k}

    \inferrule{
      \inferK{C \Cand C_x} \E \tau {k'} \\
      D = C \Cand C_x \Cand \Cleq{k'}k
    }{
      \E \vdash
      \Cleq{(\forall \kvar_i \forall (\tvar_j:k_j).\ \qual{C_x}{\tau})}{k} \Crewrite  D
    }
  \end{mathpar}

  \caption{Rewriting constraints on environments and types}
  \label{rule:typing:desugar}
\end{figure}

@@

Before presenting our constraint solving algorithms,
we place ourselves in a more general context where kinds are either variables
or constants belonging to a total bounded lattice $(\mathcal L, \lk_\Lat)$ ie.,
a lattice which admits a total order and upper and lower bounds ($l^\top$ and $l^\bot$).
We note lattice elements $l$ and $\glb_i l_i$ (resp. $\lub_i l_i$)
the greatest lower bound (resp. least upper bound) in $\mathcal L$.
%
Let $\CL$ the set of constraints in this lattice.
The full grammar of constraints is shown in \cref{grammar:constraint}.
Constraints are made of kind inequalities, conjunctions and
projections, as shown previously, along with type unification
constraints. Since types might contain kinds (for instance, on the arrows),
type unification is oriented and noted $\leq$.

@@

This algorithm is complete and computes principal normal forms.
It can be completed by a set of ``best-effort'' simplification
rules which help reduce the size of inferred signatures
\citep{DBLP:conf/popl/PottierS02,DBLP:conf/aplas/Simonet03}.
In particular, our implementation and all the signatures presented in
\cref{motivation} use a variance-based simplification
where all covariants variables are replaced by their upper bounds
and contra-variant variables are replaced by their lower bounds.

We now show that this algorithms is well-behaved with respect to entailment
and computes unique normal forms.

>>>>

An important innovation of \affe is its principal type inference.
We now formulate our inference algorithm
based on the $HM(X)$ framework~\citep{DBLP:journals/tapos/OderskySW99}.
$HM(X)$ presents a Hindley-Milner type system for a language
with qualified types where constraints are expressed in an arbitrary
constraint $X$. In particular, given some properties on the constraint system $X$,
$HM(X)$ ensures principal type inference.
We apply HM(X) to a concrete constraint language which we name $\CL$.
We adapt and extend its existing rules to support kind inference,
track linearity and handle borrows and regions. Finally, we
formulate an appropriate constraint solving algorithm for $\CL$.
In addition, our constraint solving algorithm simplifies constraints in a
principled way.
We first present various preliminary definitions
before showing our kind and type inference judgements
and our constraint solving algorithm.

@@

@@

Before presenting our constraint solving algorithm,
we place ourselves in a more general context where kinds are either variables
or constants belonging to a total bounded lattice $(\mathcal L, \lk_\Lat)$ ie.,
a lattice which admits a total order and upper and lower bounds ($l^\top$ and $l^\bot$).
We note lattice elements $l$ and $\glb_i l_i$ (resp. $\lub_i l_i$)
the greatest lower bound (resp. least upper bound) in $\mathcal L$.
%
Let $\CL$ the set of constraints in this lattice.
The full grammar of constraints is shown in \cref{grammar:constraint}.
Constraints are made of kind inequalities, conjunctions and
projections, as shown previously, along with type unification
constraints. Since types might contain kinds (for instance, on the arrows),
type unification is oriented and noted $\leq$.

@@

Our algorithm is complete, computes principal normal forms,
and already simplifies constraints significantly
(notably thanks to steps 6, 7 and 8).
It can be easily extended with further simplification phases.
In particular, our implementation and all the signatures presented in
\cref{motivation} use a variance-based simplification
where all covariant (resp. contravariant) variables are replaced by their
lower (resp. upper) bounds.
Remarkably, all the simplification mechanisms presented
here (including the variance-based one) are complete.
It is also possible to add ``best-effort'' simplification
rules which help reduce the size of inferred signatures even further
\citep{DBLP:conf/aplas/Simonet03}.
%
We now show that this algorithms is well-behaved with respect to entailment
and computes unique normal forms.
